Code used to generate paper results : Database comparison and CT characteristics

### 2024-05-14 Paper_code-Database_Comparison.Rmd
Reorganisation of paper code (in prep of adding Axelle code) and fig in pdf

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(Vennerable)
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(UpSetR)
library(ComplexHeatmap)
library(circlize)
library(SingleCellExperiment)
library(org.Hs.eg.db)
library(clusterProfiler)
library(msigdbr)
library(DOSE)
library(BiocParallel)
library(patchwork)
library(Biostrings)
```

@@

```{r load_data, warning=FALSE, message=FALSE}
# CCLE_data <- CTdata::CCLE_data()
# TCGA_TPM <- CTdata::TCGA_TPM()
# testis_sce <- CTdata::testis_sce()
####

load("../CTdata/eh_data/CCLE_data.rda")
load("../CTdata/eh_data/TCGA_TPM.rda")
load("../CTdata/eh_data/testis_sce.rda")
load("../CTdata/eh_data/mean_methylation_in_tissues.rda")

load("../CTdata/eh_data/all_genes.rda")
load("../CTdata/eh_data/HPA_cell_type_specificities.rda")
load("../CTdata/eh_data/CT_genes.rda")

CT_and_CTP_genes <- CT_genes
CT_genes <- (filter(CT_genes, CT_gene_type == "CT_gene"))
```

Common figures parameters

```{r fig_param}
legends_param <- list(
  labels_gp = gpar(col = "black", fontsize = 6),
  title_gp = gpar(col = "black", fontsize = 6),
  row_names_gp = gpar(fontsize = 4),
  annotation_name_side = "left")

legend_colors <- c("#5E4FA2", "#3288BD", "#66C2A5", "#ABDDA4", "#E6F598",
                   "#FFFFBF", "#FEE08B", "#FDAE61", "#F46D43", "#D53E4F",
                   "#9E0142")
chr_colors <- c("X-linked" = "deeppink", "Not X" = "royalblue1")

meth_colors <- c("Methylation" = "lightgreen", "Not methylation" = "gray")
```

# Pipeline figure values

Here is the code used to have all the values in the pipeline figure

```{r pipeline figures value}
table(all_genes$GTEX_category)
table(all_genes$multimapping_analysis)

table(all_genes$testis_specificity)
specific_to_pref <- all_genes %>%
  filter(GTEX_category == "testis_specific" |
           multimapping_analysis == "testis_specific") %>% 
  filter(testis_specificity == "testis_preferential")
table(specific_to_pref$not_detected_in_somatic_HPA)
table(specific_to_pref$CCLE_category)
table(specific_to_pref$TCGA_category)

all_genes %>% 
  filter(testis_specificity != "not_testis_specific") %>% 
  filter(CCLE_category == "activated" & 
           (TCGA_category == "multimapping_issue" | 
              TCGA_category == "activated")) %>% 
  pull(testis_specificity) %>% 
  table()

table(all_genes$CT_gene_type) #Before and after TSS verif
```

@@

@@

upset_CTdatabase <- fromList(CTdatabase_lost_upset)
upset(upset_CTdatabase)

 (nrow(CTdatabase_lost) - table(CTdatabase_lost$testis_specificity)[["testis_specific"]])/nrow(CTdatabase_lost)*100
```

@@

@@

upset_Jamin <- fromList(Jamin_lost_upset)
upset(upset_Jamin)
```

@@

upset_Wang <- fromList(Wang_lost_upset)
upset(upset_Wang)
```

@@

upset_Carter <- fromList(Carter_lost_upset)
upset(upset_Carter)
```

@@

upset_Bruggeman <- fromList(Bruggeman_lost_upset)
upset(upset_Bruggeman)
```

@@

@@

@@

# CT genes characteristics

```{r basic}
table(CT_genes$testis_specificity)
table(CT_genes$transcript_biotype)
```

Most genes are testis specific 
(`r table(CT_genes$testis_specificity)[["testis_specific"]]/dim(CT_genes)[1]*100`%).
Most genes are mainly protein coding genes 
(`r table(CT_genes$transcript_biotype)["protein_coding"]/dim(CT_genes)[1]*100`%).

## X chromosome and regulated by methylation

In CTexploreR, genes have been characterised as regulated by methylation or not.

```{r X_meth}
table(CT_genes$X_linked)
table(CT_genes$regulated_by_methylation)
table(CT_genes$X_linked, CT_genes$testis_specificity)
table(CT_genes$X_linked, CT_genes$regulated_by_methylation)
```

Genes are enriched on the X chromosome 
(`r table(CT_genes$X_linked)[[2]]/dim(CT_genes)[1]*100`%). Also, 
`r table(CT_genes$regulated_by_methylation)[[2]]` genes have been flagged as 
regulated by methylation 
(`r table(CT_genes$regulated_by_methylation)[[2]]/dim(CT_genes)[1]*100`%). 
It is interesting to study the link between these two characteristics.

On the chromosome X, there is a clear enrichment of CT genes regulated by 
methylation (`r table(CT_genes$X_linked, CT_genes$regulated_by_methylation)[2, 2]`/`r table(CT_genes$X_linked)[[2]]` 
chrX genes or `r table(CT_genes$X_linked, CT_genes$regulated_by_methylation)[2, 2]`/`r table(CT_genes$regulated_by_methylation)[[2]]` 
genes regulated by methylation).

Let's check that with a statistical test, I here want to see if, on there is an 
enrichment of genes regulated by methylation on the X chromosome. I need to do a
Pearson Chi square test (to know if observed proportion differ from expected 
proportion). It is a statistical method to determine if two categorical variables
have a significant correlation between them. I can directly put a matrix 
(my table) in the function

```{r chisq}
chisq.test(table(CT_genes$X_linked, CT_genes$regulated_by_methylation))
```

```{r X_meth_fig}
CT_genes$chr_factor <- factor(CT_genes$chr,
                       levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9",
                                  "10", "11", "12", "13", "14", "15", "16", "17", 
                                  "18", "19", "20", "21", "22", "X", "Y"))

CT_genes %>% 
  mutate(regulated_by_methylation = ifelse(regulated_by_methylation,
                                           "Regulated by methylation",
                                           "Not regulated by methylation")) %>% 
  ggplot(aes(x = chr_factor, fill = X_linked)) +
  geom_bar(stat = 'count') +
  scale_fill_manual(values = c("royalblue1", "deeppink")) +
  facet_wrap(~ regulated_by_methylation, ncol = 2) +
  theme_bw() +
  xlab("Chromosome") +
  ylab("Number of genes") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "none",
        axis.title = element_text(size = 10, color = "gray10"))
```

There is indeed a significative link between regulation by methylation and being
on the X chromosome. There is thus an enrichment of CT genes regulated by 
methylation on the X chromosome (and inversely).

## Tumour and methylation

Using CTexploreR functions, we can explore all CT genes or focus on some 
potential targets.

### All CT

For these heatmaps, the code comes from the function but has been copies to add 
some annotations.

```{r common_annot}
chr_mat <- as.matrix(CT_genes$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes$external_gene_name
row_ha_chr <- rowAnnotation(chr_factor = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr_factor = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes$regulated_by_methylation, CT_genes$X_linked)
```

```{r TCGA_heatmap, message = FALSE}
database <- TCGA_TPM
database$tumor <- sub(pattern = 'TCGA-', x = database$project_id, '')
database$type <- "Tumor"
database$type[database$shortLetterCode == "NT"] <- "Peritumoral"
database <- database[, order(database$tumor, database$type)]

genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]
database <- database[, database$type == "Tumor"]

column_ha_tumor <- HeatmapAnnotation(
  Tumor = database$tumor,
  border = TRUE,
  col = list(Tumor = c("BRCA" = "midnightblue", "COAD" = "darkorchid2",
                 "ESCA" = "gold", "HNSC" = "deeppink2",
                 "LUAD" = "seagreen", "LUSC" = "seagreen3",
                 "SKCM" = "red3")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

split_by <- factor(database$tumor)
col_annot <- column_ha_tumor

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

Heatmap(mat[, , drop = FALSE],
        name = "logTPM",
        column_title = paste0("Expression in TCGA samples (all)"),
        column_split = split_by,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                         legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = col_annot,
        left_annotation = left_annot)

```{r CCLE_heatmap}
database <- CCLE_data
database$type <- tolower(database$type)
genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

df_col <- data.frame("cell_line" = colData(database)$cell_line_name,
                     "type" = as.factor(colData(database)$type))
rownames(df_col) <- rownames(colData(database))
df_col <- df_col[order(df_col$type), ]

column_ha_type <- HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param,
  col = list(type = c("lung" = "seagreen3", "skin" = "red3",
                 "bile_duct" = "mediumpurple1", "bladder" = "mistyrose2",
                 "colorectal" = "plum", "lymphoma" = "steelblue1",
                 "uterine" = "darkorange4", "myeloma" = "turquoise3", 
                 "kidney" = "thistle4",
                 "pancreatic" = "darkmagenta", "brain" = "palegreen2",
                 "gastric" = "wheat3", "breast" = "midnightblue",
                 "bone" = "sienna1", "head_and_neck" = "deeppink2",
                 "ovarian" = "tan3", "sarcoma" = "lightcoral",
                 "leukemia" = "steelblue4", "esophageal"= "khaki",
                 "neuroblastoma" = "olivedrab1")))

Heatmap(mat[, rownames(df_col), drop = FALSE],
        name = "logTPM",
        column_title = "Gene Expression in tumor cell lines (CCLE)",
        column_split = factor(df_col$type),
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                          legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = c(column_ha_type),
        left_annotation = left_annot)
```

Next graph was removed from paper

```{r all_meth}
genes <- CT_genes$external_gene_name
database <- mean_methylation_in_tissues[rownames(mean_methylation_in_tissues) %in% genes]

mat <- na.omit(assay(database))
clustering_option <- TRUE

row_ha_chr_meth <- rowAnnotation(chr = chr_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

row_ha_reg_meth <- rowAnnotation(regulation = regulation_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot_meth <- c(row_ha_chr_meth, row_ha_reg_meth, gap = unit(1, "mm"))

split_meth <- data.frame(filter(CT_genes, external_gene_name %in% rownames(mat))$regulated_by_methylation, 
                    filter(CT_genes, external_gene_name %in% rownames(mat))$X_linked)

Heatmap(mat,
        column_title = 'Promoter mean methylation level by tissue',
        name = 'Meth',
        col = colorRamp2(c(1:100),
                         colorRampPalette(c("moccasin","dodgerblue4"))(100)),
        na_col = "gray80",
        cluster_rows = clustering_option,
        cluster_columns = FALSE,
        row_split = split_meth,
        row_title_gp = gpar(fontsize = 0),
        show_row_names = TRUE,
        show_heatmap_legend = TRUE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 3),
        column_names_gp = gpar(fontsize = 8),
        column_names_side = "bottom",
        row_names_side = "right",
        left_annotation = left_annot_meth)

### MAGE genes

```{r MAGE_TCGA}
MAGE_genes <- filter(CT_genes, family == "MAGE")$external_gene_name

TCGA_expression(tumor = "all", 
                genes = MAGE_genes,
                units = "log_TPM")

CCLE_expression(genes = MAGE_genes,
                 type = c("lung", "skin", "bile_duct", "bladder", 
                          "colorectal", "lymphoma", "uterine",
                          "myeloma", "kidney", "pancreatic", "brain", 
                          "gastric", "breast", "bone", "head_and_neck", 
                          "ovarian", "sarcoma", "leukemia", "esophageal",
                          "neuroblastoma"), units = "log_TPM")

normal_tissues_mean_methylation(MAGE_genes, na.omit = TRUE)

## Single Cell analysis : timing of expression

In CTexploreR, we there is single cell data from the testis, we thus can 
analyse CT genes expression during spermatogenesis.

There is only data for `r sum(CT_genes$external_gene_name %in% unique(rownames(testis_sce)))` of our `r nrow(CT_genes)` CT genes.

NB : SSC, Spermatogonia and early spermatocytes are *premeiotic cells*. Late 
spermatocytes (between both meiosis), round spermatid, elongated spermatid and 
sperm are *postmeiotic cells*.

```{r repartition_sc}
genes_avail <- 
  CT_genes$external_gene_name[CT_genes$external_gene_name %in% unique(rownames(testis_sce))]
table(CT_genes$testis_cell_type)
table(CT_genes$testis_cell_type)/length(genes_avail)*100
```

`r sum(table(CT_genes$testis_cell_type)[c("Early_spermatocyte", "Spermatogonia", "SSC")])/length(genes_avail)*100`
% of genes are mainly expressed pre-meioticly.

```{r sc_heatmap, message = FALSE}
germ_cells <- c("SSC", "Spermatogonia", "Early_spermatocyte",
                "Late_spermatocyte","Round_spermatid", "Elongated_spermatid",
                "Sperm1", "Sperm2")
somatic_cells <- c("Macrophage", "Endothelial", "Myoid", "Sertoli", "Leydig")

testis_sce_CT <- testis_sce[genes_avail, ]

mat <- SingleCellExperiment::logcounts(testis_sce_CT)

df_col <- data.frame(clusters = colData(testis_sce_CT)$clusters,
                     type = colData(testis_sce_CT)$type,
                     Donor = colData(testis_sce_CT)$Donor)
rownames(df_col) <- colnames(testis_sce_CT)

df_col <- df_col[order(df_col$type),]
df_col$lineage <- "Germ cells"
df_col$lineage[df_col$type %in% somatic_cells] <- "Somatic cells"

column_ha_type = HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  col = list(type = c("SSC" = "floralwhite", "Spermatogonia" = "moccasin",
                   "Early_spermatocyte" = "gold", 
                   "Late_spermatocyte" = "orange",
                   "Round_spermatid" = "red2", 
                   "Elongated_spermatid" = "darkred",
                   "Sperm1" = "violet", "Sperm2" = "purple", 
                   "Sertoli" = "gray", 
                   "Leydig" = "cadetblue2", "Myoid" = "springgreen3", 
                   "Macrophage" = "gray10",
                   "Endothelial" = "steelblue")),

  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

column_ha_lineage = HeatmapAnnotation(
  lineage = df_col$lineage,
  border = TRUE,
  col = list(lineage = c("Germ cells" = "salmon", "Somatic cells" = "cyan4")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

scale_lims <- c(0, quantile(rowMax(mat), 0.75))
top_annot <- c(column_ha_lineage, column_ha_type)

# Until here is what's in the function, hereunder is my addition/change in Heatmap()

CT_genes_avail <- filter(CT_genes, external_gene_name %in% genes_avail)

chr_mat <- as.matrix(CT_genes_avail$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes_avail$external_gene_name
row_ha_chr <- rowAnnotation(chr = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes_avail$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes_avail$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes_avail$regulated_by_methylation, CT_genes_avail$X_linked)

Heatmap(mat[genes_avail, rownames(df_col), drop = FALSE],
        name = "logCounts",
        column_title = "Expression in testis cells (scRNAseq)",
        column_split = df_col$type,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        show_column_names = FALSE,
        show_column_dend = FALSE,
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        col = colorRamp2(seq(scale_lims[1], scale_lims[2], length = 11),                 
                         legend_colors),
        top_annotation = top_annot,
        left_annotation = left_annot,
        heatmap_legend_param = legends_param)

## Gene function

All CT genes function

```{r GO_ORA}
msigdbr(species = "Homo sapiens" , category = "C5") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

msigdbr(species = "Homo sapiens" , category = "H") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

go_ora <- enrichGO(gene = CT_genes$ensembl_gene_id,
                   keyType = "ENSEMBL",
                   OrgDb = org.Hs.eg.db,
                   ont = "all",
                   readable = TRUE)
as_tibble(go_ora)

as_tibble(go_ora) %>% 
  arrange(desc(Count)) %>% 
  head(12) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186)) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 4)

ora_to_plot <- as_tibble(simplify(go_ora))

ora_to_plot <- ora_to_plot %>% 
  arrange(desc(Count)) %>% 
  head(9) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/102,
                           ONTOLOGY == "MF"~ Count/186))

ora_to_plot %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 3.7)
```

As we can see here, most of genes are indeed linked to functions from reproduction.
I represented here the 12 categories with the most genes, all from biological 
processes. However, they are enriched in `r nrow(as_tibble(go_ora))` different GO terms.

Is there a difference between meth reg or not ?

```{r GO_ORA_meth}
go_ora_meth <- enrichGO(gene = 
                          filter(CT_genes, regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_meth <- simplify(go_ora_meth)
as_tibble(go_ora_meth)

go_ora_not_meth <- enrichGO(gene = 
                          filter(CT_genes, !regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_not_meth <- simplify(go_ora_not_meth)
as_tibble(go_ora_not_meth)

go_ora_meth_plot <- as_tibble(go_ora_meth) %>% 
  arrange(desc(Count)) %>% 
  head(10) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/73,
                           ONTOLOGY == "MF"~ Count/80,
                           ONTOLOGY == "CC"~ Count/78)) %>%  
  mutate(regulation = "Methylation")

go_ora_not_meth_plot <- as_tibble(go_ora_not_meth) %>% 
  arrange(desc(Count)) %>% 
  head(8) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/29,
                           ONTOLOGY == "MF"~ Count/30))%>%  
  mutate(regulation = "Not methylation")

rbind(go_ora_meth_plot, go_ora_not_meth_plot) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  facet_wrap(~ regulation) + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 6, color = "gray10"),
        axis.title.x = element_text(size = 10, color = "gray10"),
        axis.title.y = element_blank())+ 
  geom_text(aes(0, y = Description, label = ID),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 2)

We've also added a column tumor suppressor and oncogene in CTexploreR. 
These information come from Cancermine.

```{r C_TS}
table(CT_genes$oncogene) 
table(CT_genes$tumor_suppressor)

table(CT_genes$oncogene, CT_genes$tumor_suppressor) 
```

>>>>

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(Vennerable)
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(UpSetR)
```

@@

```{r load_data, warning=FALSE, message=FALSE}
load("../CTdata/eh_data/all_genes.rda")
load("../CTdata/eh_data/CT_genes.rda")

CT_and_CTP_genes <- CT_genes
CT_genes <- (filter(CT_genes, CT_gene_type == "CT_gene"))

upset_text_size <- c(1.5, 1.5, 1.5, 1.5, 1.8, 2)
 #c(intersection size title, intersection size tick labels, set size title, 
 # set size tick labels, set names, numbers above bars)
```

@@

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_CTdatabase.pdf"))
plot(CTdatabase_ours, gp = gp)
dev.off()
```

@@

upset_CTdatabase <- fromList(CTdatabase_lost_upset)

upset(upset_CTdatabase,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_CTdatabase_lost.pdf"))
upset(upset_CTdatabase,
      text.scale = upset_text_size)
dev.off()
```

@@

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_omics.pdf"))
upset(upset_omics)
dev.off()
```

@@

upset_Jamin <- fromList(Jamin_lost_upset)
upset(upset_Jamin,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Jamin.pdf"))
plot(core_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Jamin_lost.pdf"))
upset(upset_Jamin,
      text.scale = upset_text_size)
dev.off()
```

@@

upset_Wang <- fromList(Wang_lost_upset)
upset(upset_Wang,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Wang.pdf"))
plot(Wang_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Wang_lost.pdf"))
upset(upset_Wang,
      text.scale = upset_text_size)
dev.off()
```

@@

upset_Carter <- fromList(Carter_lost_upset)
upset(upset_Carter,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Carter.pdf"))
plot(Carter_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Carter_lost.pdf"))
upset(upset_Carter,
      text.scale = upset_text_size)
dev.off()
```

@@

upset_Bruggeman <- fromList(Bruggeman_lost_upset)
upset(upset_Bruggeman,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Bruggeman.pdf"))
plot(Bruggeman_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Bruggeman_lost.pdf"))
upset(upset_Bruggeman,
      text.scale = upset_text_size)
dev.off()
```

@@

all_lost_upset <- 
  list(`Not testis specific` = 
         filter(lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_all <- fromList(all_lost_upset)
upset(upset_all,
      text.scale = upset_text_size)

@@

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_all_lost.pdf"))
upset(upset_all,
      text.scale = upset_text_size)
dev.off()
```

@@
